---
title: "HousePricePredictionModel"
author: "Rilind Asllani, Aaron Barnes, Brett Jenkins, Nick Russell"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# This installs the packages if you don't have them, and loads them into your environment.
# If you need more packages, add them to the packages vector.

packages <- c("tidyverse", "faux", "DataExplorer", "randomForest", "caret", "corrplot", "modelr", "stats", "rpart.plot", "mlbench")

package.check <- lapply(
  packages,
  FUN = function(x) {
    if(!require(x, character.only = TRUE)){
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

```

# House Price Prediction Model

In this source code we import, clean, and impute the data, and then we build a linear model to explain and predict future data. Finally, we output the predictions of our model to be evaluate within Kaggle.

## Custom Functions

In order to primarily evaluate the models, we creates custom functions to see the R2, RMSE, and RMSLE of the models.

```{r cars}
# Define our RMSE and RMSLE functions to evaluate our model
r2 <- function(actual, predicted){
  TSS <- sum((actual - mean(actual))^2)
  RSS <- sum((actual - predicted)^2)
  1 - RSS/TSS
}

rmse <- function(actual, fitted){
  sqrt(mean((actual - fitted)^2))
}

rmsle <- function(actual, fitted) {
  sqrt(mean((log(fitted+1) - log(actual+1))^2))
}

calc_mode <- function(x){
  # List the distinct / unique values
  distinct_values <- na.omit(unique(x))
  # Count the occurrence of each distinct value
  distinct_tabulate <- tabulate(match(x, distinct_values))
  # Return the value with the highest occurrence
  distinct_values[which.max(distinct_tabulate)]
}

```

## Cleaning the data

Cleaning the data entails understanding how much of our data is reliable, and imputing on what is not available. We then manipulate any non-normalized data to make it linear for use in a linear model.

```{r pressure, echo=FALSE}
cleaner <- function(dirty_data) {
  clean_data <- dirty_data %>%
    mutate(
      YrBltAndRemod = YearBuilt + YearRemodAdd,
      
      MSSubClass = factor(MSSubClass, level = c(20, 30, 40, 45, 50, 60, 70, 75, 80, 85, 90, 120, 150, 160, 180, 190)),
      MSZoning = MSZoning %>% replace_na("NoZone"), # Train data: No NAs; Test data: HAS NAs
      MSZoning = factor(MSZoning, levels = c("A", "C (all)", "FV", "I", "RH", "RL", "RP", "RM", "NoZone")),
      
      LotFrontage = LotFrontage %>% replace_na(0), # Train data: HAS NAs; Test data: HAS NAs
      LotFrontage = log(LotFrontage + 1),
      LotArea = log(LotArea +1),
      
      Street = factor(Street),
      Alley = Alley %>% replace_na("NoAcc"), # Replace NAs with NoAcc # HAS NAs
      Alley = factor(Alley),
      LotShape = factor(LotShape),
      LandContour = factor(LandContour),
      Utilities = Utilities %>% replace_na("Unknown"),
      Utilities = factor(Utilities, levels = c("AllPub", "NoSewr", "NoSeWa", "ELO", "Unknown")),
      LotConfig = factor(LotConfig),
      LandSlope = factor(LandSlope),
      Neighborhood = factor(Neighborhood),
      Condition1 = factor(Condition1),
      Condition2 = factor(Condition2),
      BldgType = factor(BldgType),
      HouseStyle = factor(HouseStyle),
      OverallQual = factor(OverallQual),
      OverallCond = factor(OverallCond),
      
      #These should not be factors
      #   YearBuilt = factor(YearBuilt, levels = factor(YearBuilt) %>% levels),
      #   YearRemodAdd = factor(YearRemodAdd, levels = factor(YearRemodAdd) %>% levels),
      
      RoofStyle = factor(RoofStyle),
      RoofMatl = factor(RoofMatl),
      
      Exterior1st = factor(Exterior1st), # Train data: No NAs; Test data: HAS 1 NA
      #Replacing NA with mode
      Exterior1st = Exterior1st %>% replace_na(calc_mode(levels(Exterior1st))),
      
      Exterior2nd = Exterior2nd %>% replace_na("Other"), #Train data: No NAs; Test data: HAS 1 NA
      Exterior2nd = factor(Exterior2nd),
      MasVnrType = MasVnrType %>% replace_na("None"), # Train data: HAS NAs; Test data: HAS NAs
      MasVnrType = factor(MasVnrType),
      MasVnrArea = MasVnrArea %>% replace_na(0), # Train data: HAS NAs; Test data: No NAs
      # MasVnrArea - log(MasVnrArea), # No benefit in logging it
      
      
      ExterQual = factor(ExterQual, levels = c("Po", "Fa", "TA", "Gd", "Ex")),
      ExterCond = factor(ExterCond, levels = c("Po", "Fa", "TA", "Gd", "Ex")),
      Foundation = factor(Foundation),
      BsmtQual = BsmtQual %>% replace_na("NoBsmt"), # Train data: HAS NAs; Test data: HAS NAs
      BsmtQual = factor(BsmtQual, levels = c("NoBsmt", "Po", "Fa", "TA", "Gd", "Ex")), # HAS NAs
      BsmtCond = BsmtCond %>% replace_na("NoBsmt"), # HAS NAs
      BsmtCond = factor(BsmtCond, levels = c("NoBsmt", "Po", "Fa", "TA", "Gd", "Ex")), # HAS NAs
      BsmtExposure = BsmtExposure %>% replace_na("NoBsmt"), # HAS NAs
      BsmtExposure = factor(BsmtExposure, levels = c("NoBsmt", "No", "Mn", "Av", "Gd")), # HAS NAs
      BsmtFinType1 = BsmtFinType1 %>% replace_na("NoBsmt"), # HAS NAs
      BsmtFinType1 = factor(BsmtFinType1, levels = c("NoBsmt", "Unf", "LwQ", "Rec", "BLQ", "ALQ", "GLQ")), # HAS NAs
      BsmtFinType2 = BsmtFinType2 %>% replace_na("NoBsmt"), # HAS NAs
      BsmtFinType2 = factor(BsmtFinType2, levels = c("NoBsmt", "Unf", "LwQ", "Rec", "BLQ", "ALQ", "GLQ")),
      
      # Replacing NA with 0
      BsmtFinSF1 = BsmtFinSF1 %>% replace_na(0), # Train data: No NAs, Test data: HAS 1 NA
      BsmtFinSF2 = BsmtFinSF2 %>% replace_na(0), # Test data: HAS 1 NA
      BsmtUnfSF = BsmtUnfSF %>% replace_na(0), # Test data: HAS 1 NA
      TotalBsmtSF = TotalBsmtSF %>% replace_na(0), # Test data: HAS 1 NA
      HasBsmt = ifelse(TotalBsmtSF > 0, 1, 0),
      
      X1stFlrSF = X1stFlrSF %>% replace_na(0),
      X2ndFlrSF = X2ndFlrSF %>% replace_na(0),
      
      TotalFinSF = (BsmtFinSF1 + BsmtFinSF2 + X1stFlrSF + X2ndFlrSF),
      Has2ndFloor = ifelse(X2ndFlrSF > 0, 1, 0),

      BsmtFinSF1 = log(BsmtFinSF1 +1), # Highly skewed right with most values at 0
      BsmtFinSF2 = log(BsmtFinSF2 +1), # Most values at 0
      #BsmtUnfSF = log(BsmtUnfSF), # Negative beneift when logged
      TotalBsmtSF = log(TotalBsmtSF +1), 
      TotalFinSF = log(TotalFinSF + 1),
      #   X1stFlrSF = log(X1stFlrSF), # No benefit in logging
      #   X2ndFlrSF = log(X2ndFlrSF +1),
      LowQualFinSF = log(LowQualFinSF +1),
      GrLivArea = log(GrLivArea +1),
      
      Heating = factor(Heating),
      HeatingQC = factor(HeatingQC, levels = c("Po", "Fa", "TA", "Gd", "Ex")),
      CentralAir = factor(CentralAir),
      Electrical = Electrical %>% replace_na("Mix"), # Train data: HAS NA; Test data: No NAs
      Electrical = factor(Electrical), # HAS NA
      
      FullBath = FullBath %>% replace_na(0),
      HalfBath = HalfBath %>% replace_na(0),
      BsmtFullBath  = BsmtFullBath %>% replace_na(0), # Test data: HAS 2 NAs
      BsmtHalfBath  = BsmtHalfBath %>% replace_na(0), # Test data: HAS 2 NAs
      Total_Bathrooms = (FullBath + (.5 * HalfBath) + BsmtFullBath + (.5 * BsmtHalfBath)),
      
      #    BedroomAbvGr = factor(), # Probably doesn't need any mutations
      #    KitchenAbvGr = factor(), # Probably doesn't need any mutations
      
      KitchenQual = KitchenQual %>% replace_na("TA"), # Test data: has 1 NA
      KitchenQual = factor(KitchenQual, levels = c("Po", "Fa", "TA", "Gd", "Ex")),
      
      #    KitchenQual = as.numeric(KQ_factor), #might need to be as.numeric(KitchenQual)
      #    TotRmsAbvGrd = factor(), # Probably doesn't need any mutations
      
      Functional = Functional %>% replace_na("Typ"), # Test data: has 2 NAs
      Functional = factor(Functional, levels = c("Sal", "Sev", "Maj2", "Maj1", "Mod", "Min2", "Min1", "Typ")),
      
      #    Fireplaces = factor(), # Probably doesn't need any mutations
      FireplaceQu = FireplaceQu %>% replace_na("NoFp"), # NAs should specify no fireplace # HAS NAs
      HasFireplace = ifelse(Fireplaces > 0, 1, 0),
      FireplaceQu = factor(FireplaceQu, levels = c("NoFp", "Po", "Fa", "TA", "Gd", "Ex")),  # HAS NAs
      
      GarageType = GarageType %>% replace_na("NoGrge"), #NAs should specify no garage # HAS NAs
      GarageType = factor(GarageType), # HAS NAs
      GarageYrBlt = GarageYrBlt %>% replace_na(0), # NAs should use the median [1980] to not mess with the data # HAS NAs
      GarageYrBlt = log(GarageYrBlt +1),
      GarageFinish = GarageFinish  %>% replace_na("NoGrge"), # HAS NAs
      GarageFinish = factor(GarageFinish), # NAs should specify no garage # HAS NAs
      GarageCars = GarageCars %>% replace_na(0), # Test data: HAS 1 NA
      GarageArea = GarageArea %>% replace_na(0), # Test data: HAS 1 NA
      HasGarage = ifelse(GarageArea > 0, 1, 0),
      #   GarageArea = log(GarageArea),
      GarageQual = GarageQual %>% replace_na("NoGrge"), # NAs should specify no garage # HAS NAs
      GarageQual = factor(GarageQual, levels = c("NoGrge", "Po", "Fa", "TA", "Gd", "Ex")), # HAS NAs
      GarageCond = GarageCond %>% replace_na("NoGrge"), # NAs should specify no garage # HAS NAs
      GarageCond = factor(GarageCond, levels = c("NoGrge", "Po", "Fa", "TA", "Gd", "Ex")), # HAS NAs
      
      PavedDrive = factor(PavedDrive),
      
      Total_Porch_SF = (OpenPorchSF + X3SsnPorch + EnclosedPorch + ScreenPorch + WoodDeckSF),
      OpenPorchSF = log(OpenPorchSF +1),
      X3SsnPorch = log(X3SsnPorch +1),
      EnclosedPorch = log(EnclosedPorch +1),
      #   ScreenPorch = log(ScreenPorch), # No benefit in logging
      #   WoodDeckSF = log(WoodDeckSF), # No beneift in logging
     
      #   PoolArea = log(PoolArea), # No benfit in logging
      HasPool = ifelse(PoolArea > 0, 1, 0),
      PoolQC = PoolQC %>% replace_na("NoPool"), #NAs should specify no pool # HAS NAs
      PoolQC = factor(PoolQC, levels = c("NoPool", "Fa", "TA", "Gd", "Ex")), # HAS NAs
      
      Fence = Fence %>% replace_na("NoFnc"), #NAs should specify no fence # HAS NAs
      Fence = factor(Fence), # HAS NAs
      
      MiscFeature = MiscFeature %>% replace_na("NoMiscF"), #NAs should specify no amenities # HAS NAs
      MiscFeature = factor(MiscFeature), # HAS NAs
      #    MiscVal = factor(), # Probably doesn't need any mutations

      MoSold = factor(MoSold),
      #    YrSold = factor(), # Probably doesn't need any mutations
      
      SaleType = factor(SaleType),
      SaleType = SaleType %>% replace_na(calc_mode(levels(SaleType))), # Test data has 1 NA
      
      SaleCondition = factor(SaleCondition)
    )
  return(clean_data)
}

```

## Read data into environment

Now that we have definied a function to clean this data, we will import the data, run it through the cleaner function, and save it to a variable.

```{r}
# Read in the data from CSV to RStudio
train_data <- read.csv("train.csv") %>%
  cleaner() %>%
  mutate(SalePrice = log(SalePrice))

# Let's verify that we don't have any missing training data
complete.cases(train_data) %>% all


test_data <- read.csv("test.csv") %>%
  cleaner()

# Let's verify that we don't have any missing test data
complete.cases(test_data) %>% all

# test_data %>%
#    sapply(function(x) sum(is.na(x)))
```

## Build linear model

The data has been saved to a variable, now we will build out the linear model.

```{r message = FALSE, warning = FALSE}
# Create linear model with GLMNET using CARET
set.seed(123)
c_lm <- train(SalePrice ~
                MSSubClass +
                MSZoning +
                LotFrontage +
                LotArea +
                #   Street +
                #   Alley +
                #   LotShape +
                #   LandContour +
                #   Utilities +
                LotConfig +
                LandSlope +
                Neighborhood +
                Condition1 +
                Condition2 +
                BldgType +
                HouseStyle +
                OverallQual * Neighborhood + 
                OverallQual * GrLivArea +
                OverallCond * Neighborhood +
                YearBuilt +
                YearRemodAdd +
                RoofStyle +
                RoofMatl +
                Exterior1st +
                Exterior2nd +
                MasVnrType +
                MasVnrArea +
                ExterQual +
                ExterCond +
                Foundation +
                BsmtQual +
                BsmtCond +
                BsmtExposure +
                BsmtFinType1 +
                BsmtFinSF1 + 
                BsmtFinType2 +
                BsmtFinSF2 +
                BsmtUnfSF + 
                TotalBsmtSF +
                Heating +
                HeatingQC +
                CentralAir +
                Electrical +
                #   X1stFlrSF +  # Highly correlated to 'TotalBsmtSF'
                #   X2ndFlrSF +
                LowQualFinSF +
                GrLivArea * Neighborhood +
                BsmtFullBath +
                BsmtHalfBath +
                FullBath +
                HalfBath +
                BedroomAbvGr +
                KitchenAbvGr +
                KitchenQual +
                #   TotRmsAbvGrd +  # Highly correlated to 'GrLivArea'
                Functional +
                Fireplaces +
                FireplaceQu +
                GarageType +
                #   GarageYrBlt +
                GarageFinish +
                GarageCars +
                #   GarageArea +  # Highly correlated to 'GarageCars'
                GarageQual +
                GarageCond +
                PavedDrive +
                WoodDeckSF +
                OpenPorchSF +
                EnclosedPorch +
                X3SsnPorch +
                ScreenPorch +
                #   PoolArea +
                #   PoolQC +
                Fence +
                # MiscFeature +
                MiscVal +
                MoSold +
                YrSold +
                SaleType +
                SaleCondition +
                TotalFinSF +
                TotalFinSF * Neighborhood +
                Total_Bathrooms +
                Total_Porch_SF +
                HasPool +
                Has2ndFloor +
                HasGarage +
                HasBsmt +
                HasFireplace,
                data = train_data,
                preProcess = c("center", "scale"),
                method = "glmnet")
```

## Evaluating the linaer model

We will evaluate the model now by looking at the R2, RMSE, and other metrics that will give us an idea of how well it predicts the in-sample, and out-of-sample data.

```{r}
#In-sample performance
#R^2
r2(exp(train_data$SalePrice), exp(fitted(c_lm)))

#RSME
rmse(exp(train_data$SalePrice), exp(fitted(c_lm)))

#RSMLE
rmsle(exp(train_data$SalePrice), exp(fitted(c_lm)))

#Out-of-sample performance
best_alpha <- c_lm$results[c_lm$results$alpha==c_lm$bestTune$alpha, ]
best_alpha_lambda <- best_alpha[best_alpha$lambda==c_lm$bestTune$lambda, ]

#R^2
best_alpha_lambda$Rsquared

#RSMLE
best_alpha_lambda$RMSE

#90% Ridge Regression and 10# Lasso Regression Coefficients Removed
c_lm_coeffs <- coef(c_lm$finalModel, c_lm$bestTune$lambda)
c_lm_Rmvd_Coef <- c_lm_coeffs[c_lm_coeffs[,1]==0,0]
# c_lm_Rmvd_Coef
nrow(c_lm_Rmvd_Coef)
c_lm_Kept_Coef <- c_lm_coeffs[c_lm_coeffs[,1]!=0,0]
# c_lm_Kept_Coef
nrow(c_lm_Kept_Coef)

c_lm$bestTune
```


## Random Forest

We will also use a Random Forest model to compare how well our linear model works. If the random forst is better, we will proceed with using the Random Forest for predictions.

```{r}
set.seed(123)
c_rf <- train(SalePrice ~ .,
              data = train_data,
              method = "rf",
              preProcess = c("center", "scale"),
              importance = TRUE,
              trControl = trainControl(method = "cv", number = 3))

print(c_rf)
```

It doesn't perform as well as our linear model, so we will proceed with using the linear model for predicting outcomes.


## Creating the submission file

Now that we have a model we can rely on, we will use the model to predict the details on the test data and write those predictions to a file.

```{r}
submission_data <- test_data %>%
  select(Id) %>%
  mutate(SalePrice = exp(predict(c_lm, test_data)))

write.csv(submission_data, "submission_file.csv", row.names=FALSE)
```



```{r}
###################################################################################
################## DELETE THIS WHEN DONE WITH REPORT ##############################

#Kaggle submission results:

#All predictors, no transformations 
# Result := Log RSME 0.20247

#All predictors, log transformations on:
#   SalePrice
#   LotFrontage
#   LotArea
#   X1stFlrSF
#   GrLivArea
# Result := Log RSME 0.15104

#All predictors, scaled and centered
# Result := Log RSME 0.19524

#All predictors, scaled and centered, log transformations on:
#   SalePrice
#   LotFrontage
#   LotArea
#   X1stFlrSF
#   GrLivArea
# Result := Log RSME 0.15104

#All predictors, ridge/lasso, log transformations on:
#   SalePrice
#   LotArea
#   X1stFlrSF
#   GrLivArea
# Kaggle Result := Log RSME 0.14007
#   alpha       lambda      RMSE  Rsquared        MAE      RMSESD RsquaredSD       MAESD
#3  0.10 0.0583203499 0.1421900 0.8742627 0.08719893 0.013919797 0.02815391 0.003560103

#All predictors, ridge/lasso, scaled and centered, log transformations on:
#   SalePrice
#   LotArea
#   X1stFlrSF
#   GrLivArea
#   alpha       lambda      RMSE  Rsquared        MAE      RMSESD RsquaredSD       MAESD
#3  0.10 0.0583203499 0.1402955 0.8790267 0.08679432 0.012036779 0.02125528 0.003616432
# Kaggle Result := Log RSME 0.14007

#All predictors minus highly correlated predictors (X1stFlrSF, TotRmsAbvGrd, and GarageArea), and multiplying OverallQual * by GrLivArea
# Kaggle Result: = Log RMSE 0.13953

#Rubric score	Kaggle score (log RMSE)
# 10  <.12
# 9	  <.13
# 8	  <.14 
# 7	  <.15 <- Current Score
# 6	  <.16 
# 5	  <.17
```